Comparing Power Optimizations in the Reorder Buffer

### 2011-06-08 paper/proj.tex
Implementation section on dynamic ROB

\subsubsection{Dynamically Sized ROB}
The dynamically sized ROB is an optimization on the circular ROB where the buffer is divided into partitions, and those partitions deemed unnecessary are powered off to conserve energy. Our implementation of the basic dynamic ROB is behaviorally the same as that detailed in \cite{kucuk2}.

A conservative approach of deallocating strictly from one end of the buffer, and allocating strictly towards that end is followed. For example, if the ROB consisted of four partitions, downsizing the buffer would turn off the fourth, and then the third. At this point, if we were to increase the size of the buffer, the third partition would be turned on first, and the fourth partition next.

>>>>

\subsubsection{Dynamically Sized ROB}
Our implementation of the basic dynamic ROB is behaviorally the same as that detailed in \cite{kucuk2} and the background section.

For {\it P} partitions, instead of allocating one region of memory as in the circular ROB we allocate {\it P} regions, and associate with each region two status bits, denote the four states a ROB partition can have: {\tt ENABLED}, {\tt DISABLED}, {\tt TO\_ENABLE}, and {\tt TO\_DISABLE}. Also, where pointer incrementation in the circular buffer was done as:

\begin{center}
$ new\_ptr = (old\_ptr + 1) \: \% \: ROB\_size $
\end{center}

Which is no long accurate in the dynamic ROB. The pointer incrementation module for the dynamic ROB was updated to check if the incremented pointer is stepping into a new partition, and to step over that partition if necessary. When incrementing the head pointer, only partitions labeled {\tt ENABLED} and {\tt TO\_DISABLE} are valid. When incrementing the tail pointer, only partitions labeled {\tt ENABLED} are valid.

The entry fetch module also requires updating. Whereas in the circular buffer, an entry is located simply by taking the head or tail pointer as the offset into the buffer, the dynamic buffer must now map the pointer to a particular partition and an offset into that partition.

At cycle completion, a module was added to perform sampling and resizing. Every cycle, the module checks for a stall due to the buffer being full, and if so the {\it overflow\_count} is incremented. Every {\it sample\_period} cycles, the buffer's {\it active\_size} is sampled. Every {\it update\_period}, the buffer's size is updated by marking the appropriate partitions {\tt TO\_ENABLE} or {\tt TO\_DISABLE}. Each cycle, the head and tail pointers are checked against the marked partitions to see if their state can be updated to {\tt ENABLED} or {\tt DISABLED}.

\cite{kucuk3} takes a conservative approach of deallocating strictly from one end of the buffer, and also allocating strictly towards that end. For example, if the ROB consisted of four partitions, downsizing the buffer would turn off the fourth, and then the third. At this point, if we were to increase the size of the buffer, the third partition would be turned on first, and the fourth partition next.
